---
title: "Statistical Analysis of the ToothGrowth Data"
author: "Arun K Viswanathan"
date: "January 17, 2015"
output: pdf_document
---

# Overview

This report analyzes the ToothGrowth dataset in R.

# Summary of the Tooth Growth dataset

The _Tooth Growth_ dataset in R shows the effect of Vitamin C on tooth growth
in guinea pigs. The data provides is the length of odontoblasts (teeth) in 
each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 
2 mg) with each of two delivery methods (orange juice or ascorbic acid).

The data contains three variables:

Variable | Data Type | Description
---------|-----------|------------
`len`    | `numeric` | Tooth length
`supp`   | `factor`  | Supplement type (VC or OJ)
`dose`   | `numeric` | Dose in milligrams

The dataset contains a total of `r nrow(ToothGrowth)` observations. Here's a
summary of the raw data. 

```{r, echo=TRUE}
data <- transform(ToothGrowth, dose = as.factor(dose))
summary(data)
```

The plot below shows the tooth growth ranges for each supplement at each dosage
level.

```{r, echo=TRUE}
library(ggplot2)
g <- ggplot(data, aes(x = supp, y = len, fill = dose)) + geom_boxplot() +
    facet_grid(. ~ dose) + 
    ggtitle("Tooth Growth by Supplement Type") +
    xlab("Supplement") + ylab("Length of Tooth Growth") +
    guides(fill = guide_legend(title = "Dose"))
print(g)
```

# Tooth growth analysis by supplement

Assuming that the tooth growth data is _iid_, the 95% confidence intervals for 
tooth growth for each supplement can be computed with the code below. The plots 
show the tooth growth for each supplement as a histogram.

```{r,echo=TRUE}
conf <- lapply(split(data, data$supp), function(dataForSupp) {
    lens <- dataForSupp$len
    mean(lens) + c(-1, 1) * qnorm(0.975) * sd(lens) / sqrt(length(lens))
    })
analysis <- lapply(names(conf), function(supp) {
    lens <- data[data$supp == supp, "len"]    
    h <- hist(lens, xlim = range(data$len), ylim = range(0, 10),
              main = paste0("Histogram of tooth growth with ", supp),
              xlab = "Tooth Growth", ylab = "Frequency", col = "lightyellow2")
    xfit <- seq(min(lens), max(lens), length = 40) 
    yfit <- dnorm(xfit, mean = mean(lens), sd = sd(lens)) 
    yfit <- yfit * diff(h$mids[1:2]) * length(lens) 
    lines(xfit, yfit, col = "darkgreen", lt = 1, lw = 3)
    abline(v = mean(lens), col = "darkred", lt = 2, lw = 2)
    abline(v = conf[[supp]], col = "darkred", lt = 1, lw = 3)
    axis(1, font = 2, at = conf[[supp]], labels = sprintf("%2.2f", conf[[supp]]))
    paste0("For supplement ", supp, ", the tooth growth is between ", 
           sprintf("%2.2f", conf[[supp]][1]), " and ", 
           sprintf("%2.2f", conf[[supp]][2]), " with 95% confidence.")
    })
```

```r analysis[[1]]``` 

```r analysis[[2]]```

It can be clearly seen that orange juice (OJ) produces greater tooth growth
compared to ascorbic acid (VC).

# Tooth growth analysis by dose

Assuming that the tooth growth data is _iid_, the 95% confidence intervals for 
tooth growth for each dose can be computed with the code below. The plots 
show the tooth growth for each dose as a histogram.

```{r,echo=TRUE}
conf <- lapply(split(data, data$dose), function(dataForDose) {
    lens <- dataForDose$len
    mean(lens) + c(-1, 1) * qnorm(0.975) * sd(lens) / sqrt(length(lens))
    })
analysis <- lapply(names(conf), function(dose) {
    lens <- data[data$dose == dose, "len"]    
    h <- hist(lens, breaks = 8, xlim = range(data$len), ylim = range(0, 6.5),              
              main = paste0("Histogram of tooth growth with dose ", dose, "mg"),
              xlab = "Tooth Growth", ylab = "Frequency", col = "lightyellow2")
    xfit <- seq(min(lens), max(lens), length = 40) 
    yfit <- dnorm(xfit, mean = mean(lens), sd = sd(lens)) 
    yfit <- yfit * diff(h$mids[1:2]) * length(lens) 
    lines(xfit, yfit, col = "darkgreen", lt = 1, lw = 3)
    abline(v = mean(lens), col = "darkred", lt = 2, lw = 2)
    abline(v = conf[[dose]], col = "darkred", lt = 1, lw = 3)
    axis(1, font = 2, at = conf[[dose]], labels = sprintf("%2.2f", conf[[dose]]))
    paste0("For dose ", dose, "mg, the tooth growth is between ", 
           sprintf("%2.2f", conf[[dose]][1]), " and ", 
           sprintf("%2.2f", conf[[dose]][2]), " with 95% confidence.")
    })
```

```r analysis[[1]]``` 

```r analysis[[2]]``` 

```r analysis[[3]]```

It can be clearly seen that increasing the dose increases the tooth growth.

