---
title: "Statistical Analysis of the Exponential Distribution in R"
author: "Arun K Viswanathan"
date: "January 17, 2015"
output: pdf_document
---

# Overview

This paper investigates the exponential distribution in R and compares it with 
the Central Limit Theorem.  

# Simulations

```{r,echo=TRUE}
lambda <- 0.2     # Lamdba
n <- 40           # Number of exponentials
numsim <- 100000  # Number of simulations
```

We simulate an exponential distribution using `rexp`, with `r n` exponentials 
and lamda of `r lambda`. The simulation is performed 
`r format(numsim, scientific=FALSE)` times.  The code below shows the 
simulation in action:

```{r,echo=TRUE}
sim <- matrix(rexp(n * numsim, lambda), numsim) 
```

# Sample Mean versus Theoretical Mean

The code below computes the sample mean from the simulated data.

```{r,echo=TRUE}
mean.theoretical <- 1/lambda
means.sim <- apply(sim, 1, mean)
mean.sim <- mean(means.sim)
mean.error <- abs(mean.sim - mean.theoretical) / mean.theoretical * 100
```

The sample mean is **`r format(mean.sim)`** while the theoretical mean is 
**`r mean.theoretical`**. This is an error of **`r format(mean.error)`%**. This
clearly shows that the Central Limit Theorem in action as the mean of the
simulated means is almost the same as the theoretical mean.

# Sample Variance versus Theoretical Variance

The sample variance can be computed using the code show below:

```{r,echo=TRUE}
sd.theoretical <- 1/lambda
var.theoretical <- sd.theoretical ^ 2
sds.sim <- apply(sim, 1, sd)
sd.sim <- mean(sds.sim)
var.sim <- sd.sim ^ 2
var.error <- abs(var.sim - var.theoretical) / var.theoretical * 100
```

The average sample variance is **`r format(var.sim)`** (standard deviation of 
`r format(sd.sim)`) while the theoretical variance is 
**`r var.theoretical`** (standard deviation of `r format(sd.theoretical)`). 
This is an error of **`r format(var.error)`%** in the variance.

# Distribution

The plot below shows the distribution of the _simulated samples_.

```{r,echo=TRUE}
library(ggplot2)
g <- ggplot(data.frame(x = as.vector(sim)), aes(x = x)) +
    geom_histogram(aes(fill = ..count..), binwidth = 1) +
    ggtitle("Distribution of samples (exponential distribution)") +
    xlab("Sample Value") +
    ylab("Frequency") 
print(g)
````

The histogram below shows the distribution of the _sample means_. The red 
vertical dashed line is the mean of the sample means and is at `r mean.sim`. A
normal distribution is overlaid over the histogram in dark green.

```{r,echo=TRUE}
h <- hist(means.sim, 
          breaks = 30, 
          col = "lightyellow2",
          main = "Distribution of Sample Means",
          xlab = "Sample Mean",
          ylab = "Frequency") 
xfit <- seq(min(means.sim), max(means.sim), length = 40) 
yfit <- dnorm(xfit, mean = mean.sim, sd = sd(means.sim)) 
yfit <- yfit * diff(h$mids[1:2]) * length(means.sim) 
lines(xfit, yfit, col = "darkgreen", lt = 1, lw = 3)
abline(v = mean.sim, col = "darkred", lt = 2, lw = 3)
```

Looking at the histogram above, the distribution of the sample means appears
to be close to a normal distribution.

